# Flextable

<!--
Hidden comments placeholder
-->

```{r echo=FALSE}
knitr::include_graphics("https://picsum.photos/200/300")
```

## â­ï¸Overview

Using the Flextable package

Some overlap with:   

* officer.Rmd    
* descriptive_notes.Rmd   
* tables_descriptive.Rmd   

Good examples to check out:

* LEAD panel monthly report.   
* Sun Study report.   
* stroke study -> table_characteristics_by_network.Rmd.   
* L2C quarterly reports.    


## ðŸŒŽUseful websites

* [Using the flextable R package](https://ardata-fr.github.io/flextable-book/)   
* [Flextable R package](https://davidgohel.github.io/flextable/)


## ðŸ“¦Load packages

```{r message=FALSE}
library(dplyr, warn.conflicts = FALSE)
library(officer, warn.conflicts = FALSE)
library(flextable, warn.conflicts = FALSE)
```


## ðŸ”¢Simulate data 

```{r}
data("mtcars")
```


## Adding blank rows

Actually, might just create this in Word and post on [StackOverflow](https://stackoverflow.com/questions/64932726/add-a-blank-row-in-flextable).

When creating tables in Word reports, I often want to add blank rows in between variables. As a trivial toy example:

```{r}
doc <- read_docx()
```

```{r}
table_no_breaks <- mtcars %>% 
  count(cyl)

table_no_breaks
```

```{r}
table_no_breaks_ft <- flextable(table_no_breaks)

table_no_breaks_ft
```

```{r}
doc <- doc %>% 
  body_add_flextable(table_no_breaks_ft) %>% 
  body_add_par("")
```

```{r}
print(
  doc, 
  "examples/flextable_no_blank_rows.docx"
)
```

Results in a table that looks like this: [flextable_no_blank_rows.docx](https://www.dropbox.com/s/09weee9je1sdj7f/flextable_no_blank_rows.docx?dl=0)

I can add line breaks directly to the data frame like this:

```{r}
table_breaks <- table_no_breaks %>% 
  mutate(
    across(
      everything(),
      as.character
    )
  ) %>% 
  add_row(cyl = NA, n = NA, .after = 1) %>% 
  add_row(cyl = NA, n = NA, .after = 3) %>%
  add_row(cyl = NA, n = NA, .after = 5)

table_breaks
```

```{r}
table_breaks_ft <- flextable(table_breaks)

table_breaks_ft
```

```{r}
doc <- doc %>% 
  body_add_flextable(table_breaks_ft)
```

```{r}
print(
  doc, 
  "examples/flextable_blank_rows.docx"
)
```

Which results in the Word table that I want: [flextable_blank_rows.docx](https://www.dropbox.com/s/smq0x7qlf5muhvf/flextable_blank_rows.docx?dl=0)

However, I would prefer to be able to add the blank rows directly to the flextable rather than the data frame the flextable is built from. To me, this is a formatting issue rather than a data issue, and I prefer to keep data issues (manipulation of data frames) and formatting issues (manipulation of Word tables) separate. Any advice is greatly appreciated!

## Autofit to window

```{r}
table_breaks_ft %>% 
  set_table_properties(layout = "autofit")
```

## Change font to TNR

```{r}
table_breaks_ft %>% 
  font(fontname = "Times New Roman", part = "all")
```

## Bold part of the table title

From SO 2020-08-23: https://stackoverflow.com/questions/63530204/is-there-a-way-to-bold-part-of-a-character-string-being-passed-to-add-header-lin?noredirect=1#comment112346997_63530204

```{r}
mtcars_ft <- flextable(head(mtcars)) %>% 
  # Add a blank title line to top of table
  add_header_lines("") %>% 
  # Use compose to bold "Table #."
  compose(
    i = 1, part = "header",
    value = as_paragraph(
      as_chunk("Table 1. ", props = fp_text(bold = TRUE)),
      "Here is my example mtcars ft."
    ),
  )

mtcars_ft
```


## Conditional formatting

From LEAD panel summarize votes -- Doesn't work here. Need to turn this into a working example at some point.

```{r eval=FALSE}
title <- "Table 2. Presence/absence of unanimous agreement for each abuse type by case number."

summary_agreement_ad_abuse_type_ft <- flextable(
  # Remove unneeded columns
  summary_agreement_ad_abuse_type %>% 
    select(
      CaseID, physical, sexual, emotional, neglect, 
      abandonment, financial, selfneglect
    )
  ) %>% 
  # Column width: Trial and error
  # Make a table and play with properties
  width(
    j = c(1:8), 
    width = c(0.98, 0.66, 0.56, 0.78, 0.71, 1.01, 0.71, 0.90)
  ) %>% 
  # Improve readability of column headers
  set_header_labels(CaseID = "Case Number", selfneglect = "Self Neglect") %>% 
  # Add title to top of table
  add_header_lines(title) %>%
  # Change font to times new roman
  font(fontname = "Times New Roman", part = "all") %>% 
  # Change background color of first column
  bg(j = 1, bg = "#E5E8E8", part = "body") %>% 
  # Center column headings
  align(i = 2, align = "center", part = "header") %>% 
  # Center body text
  align(align = "center", part = "body") %>% 
  # Conditionally format disagree to red
  color(i = ~ physical == "Disagree", j = c("CaseID", "physical"), color = "red") %>%
  color(i = ~ sexual == "Disagree", j = c("CaseID", "sexual"), color = "red") %>%
  color(i = ~ emotional == "Disagree", j = c("CaseID", "emotional"), color = "red") %>%
  color(i = ~ neglect == "Disagree", j = c("CaseID", "neglect"), color = "red") %>%
  color(i = ~ abandonment == "Disagree", j = c("CaseID", "abandonment"), color = "red") %>%
  color(i = ~ financial == "Disagree", j = c("CaseID", "financial"), color = "red") %>%
  color(i = ~ selfneglect == "Disagree", j = c("CaseID", "selfneglect"), color = "red")

# For checking
# summary_agreement_ad_abuse_type_ft
```


<!--
Clean up and add session info
-->

```{r echo=FALSE}
rm(list = ls())
```

```{r echo=FALSE}
sessionInfo()
```
