# Flextable

<!--
Hidden comments placeholder
-->

```{r echo=FALSE}
knitr::include_graphics("https://picsum.photos/200/300")
```

## ⭐️Overview

Using the `flextable` package

Some overlap with:   

* [Officer]   
* [Tables]    


## 🌎Useful websites

* [Using the flextable R package](https://ardata-fr.github.io/flextable-book/)   
* [Flextable R package](https://davidgohel.github.io/flextable/)


## 📦Load packages

```{r message=FALSE}
library(dplyr, warn.conflicts = FALSE)
library(officer, warn.conflicts = FALSE)
library(flextable, warn.conflicts = FALSE)
```


## Load data 

```{r}
data("mtcars")
```


## 🔵 i and j

i = rows    
j = columns    


## 🔵 Flextable wrappers

We can wrap flextable functions that we use repeatedly in a wrapper.

```{r}
my_ft_theme <- function(ft, ...) {
  # Remove vertical cell padding
  ft <- padding(ft, padding.top = 0, padding.bottom = 0, part = "all")
  
  # Change font to TNR 11
  ft <- font(ft, fontname = "Times New Roman", part = "all")
  ft <- fontsize(ft, part = "all", size = 11)
  ft
}
```

And then use it like this:

```{r}
flextable(head(mtcars)) %>%
  my_ft_theme()
```


## 🔵 Adding/modifying table content

### Add table title with bold part

From SO 2020-08-23: https://stackoverflow.com/questions/63530204/is-there-a-way-to-bold-part-of-a-character-string-being-passed-to-add-header-lin?noredirect=1#comment112346997_63530204

```{r}
mtcars_ft <- flextable(head(mtcars)) %>% 
  # Add a blank title line to top of table
  add_header_lines("") %>% 
  # Use compose to bold "Table #."
  compose(
    i = 1, part = "header",
    value = as_paragraph(
      as_chunk("Table 1. ", props = fp_text(bold = TRUE)),
      "Here is my example mtcars ft."
    ),
  )

mtcars_ft
```

### Adding blank rows

I created a post about this on [StackOverflow](https://stackoverflow.com/questions/64932726/add-a-blank-row-in-flextable).

When creating tables in Word reports, I often want to add blank rows in between variables. As a trivial toy example:

```{r}
doc <- read_docx()
```

```{r}
table_no_breaks <- mtcars %>% 
  count(cyl)
table_no_breaks
```

```{r}
table_no_breaks_ft <- flextable(table_no_breaks)
table_no_breaks_ft
```

```{r}
doc <- doc %>% 
  body_add_flextable(table_no_breaks_ft) %>% 
  body_add_par("")
```

```{r}
print(
  doc, 
  "examples/flextable_no_blank_rows.docx"
)
```

Results in a table that looks like this: [flextable_no_blank_rows.docx](https://www.dropbox.com/s/09weee9je1sdj7f/flextable_no_blank_rows.docx?dl=0)

I can add line breaks directly to the data frame like this:

```{r}
table_breaks <- table_no_breaks %>% 
  mutate(
    across(
      everything(),
      as.character
    )
  ) %>% 
  add_row(cyl = NA, n = NA, .after = 1) %>% 
  add_row(cyl = NA, n = NA, .after = 3) %>%
  add_row(cyl = NA, n = NA, .after = 5)

table_breaks
```

```{r}
table_breaks_ft <- flextable(table_breaks)
table_breaks_ft
```

```{r}
doc <- doc %>% 
  body_add_flextable(table_breaks_ft)
```

```{r}
print(
  doc, 
  "examples/flextable_blank_rows.docx"
)
```

Which results in the Word table that I want: [flextable_blank_rows.docx](https://www.dropbox.com/s/smq0x7qlf5muhvf/flextable_blank_rows.docx?dl=0)

However, I would prefer to be able to add the blank rows directly to the flextable rather than the data frame the flextable is built from. To me, this is a formatting issue rather than a data issue, and I prefer to keep data issues (manipulation of data frames) and formatting issues (manipulation of Word tables) separate. Any advice is greatly appreciated!

### Change column header text 

Example from [Tables] chapter, which is from L2C smartphone paper.

[Compose chapter in flextable book](https://ardata-fr.github.io/flextable-book/cell-content-1.html#multi-content)

```{r}
# Calculate group n's
n_outcome <- c("No" = 53, "Yes" = 47)
```

```{r}
# Simulate data
table <- tribble(
  ~var,                ~No,                     ~Yes,
  "age",               "34.89 (30.19 - 39.58)", "35.38 (30.58 - 40.19)",
  "",                  "",                      "",
  "age_group",         "",                      "",
  "  Younger than 30", "58.49 (44.63 - 71.12)", "61.70 (46.88 - 74.63)",
  "  30 and Older",    "41.51 (28.88 - 55.37)", "38.30 (25.37 - 53.12)"
)
```

```{r}
flextable(table) %>% 
  width(width = c(3, 2, 2)) %>% 
  # Center the final two columns
  align(j = c(2, 3), align = "center", part = "all") %>% 
  # Change header names -- add subgroup Ns to headers
  set_header_labels(
    var = "Characteristic",
    No = paste0("No\n(n=", n_outcome["No"], ")"), 
    Yes = paste0("Yes\n(n=", n_outcome["Yes"], ")")
  ) %>% 
  # Bold column headers
  bold(part = "header") 
```

### Change row header text 

Example from [Tables] chapter, which is from L2C smartphone paper.

[Compose chapter in flextable book](https://ardata-fr.github.io/flextable-book/cell-content-1.html#multi-content)

```{r}
# Simulate data
table <- tribble(
  ~var,                ~No,                     ~Yes,
  "age",               "34.89 (30.19 - 39.58)", "35.38 (30.58 - 40.19)",
  "",                  "",                      "",
  "age_group",         "",                      "",
  "  Younger than 30", "58.49 (44.63 - 71.12)", "61.70 (46.88 - 74.63)",
  "  30 and Older",    "41.51 (28.88 - 55.37)", "38.30 (25.37 - 53.12)"
)
```

```{r}
flextable(table) %>% 
  width(width = c(3, 2, 2)) %>% 
  # Change text by location 
  compose(i = 1, j = 1, value = as_paragraph("Age, mean (95% CI)")) %>% 
  # Change text conditionally
  compose(i = ~ var == "age_group", j = 1, value = as_paragraph("Age group, row percent (95% CI)"))
```

### Add footnote 

Example from [Tables] chapter, which is from L2C smartphone paper.

[Footnote documentation](https://davidgohel.github.io/flextable/reference/footnote.html)

```{r}
# Simulate data
table <- tribble(
  ~var,                ~No,                     ~Yes,
  "age",               "34.89 (30.19 - 39.58)", "35.38 (30.58 - 40.19)",
  "",                  "",                      "",
  "age_group",         "",                      "",
  "  Younger than 30", "58.49 (44.63 - 71.12)", "61.70 (46.88 - 74.63)",
  "  30 and Older",    "41.51 (28.88 - 55.37)", "38.30 (25.37 - 53.12)"
)
```

Add a superscript "1" behind age and a numbered footnote at the bottom of the table. 

```{r}
flextable(table) %>% 
  width(width = c(3, 2, 2)) %>% 
  footnote(i = 1, j = 1, value = as_paragraph("Test Footnote"))
```
Or more than one at at a time:

```{r}
flextable(table) %>% 
  width(width = c(3, 2, 2)) %>% 
  footnote(
    i = c(1, 3), j = 1,
    value = as_paragraph(
      c("Age in years.", "Age grouped above and below median.")
    )
  )
```


## 🔵 Formatting 

### Change font to TNR

```{r}
flextable(head(mtcars)) %>% 
  font(fontname = "Times New Roman", part = "all")
```

### Conditional formatting

```{r}
flextable(head(mtcars)) %>% 
  # If cyl is 4 turn all text blue
  color(i = ~ cyl == 4, color = "blue") %>% 
  # mpg is greater than average mpg then format the color to red
  color(i = ~ mpg > mean(mpg), j = "mpg", color = "red")
```


## 🔵Layout 

### Autofit to contents

```{r}
flextable(head(mtcars)) %>% 
  autofit()
```

## 🔵 Examples

Other good examples to check out:

* Sun Study report.   
* stroke study -> table_characteristics_by_network.Rmd.   
* L2C quarterly reports.     
* L2C paper_smartphone_app   

### LEAD panel summarize votes

Simulate data

```{r}
summary_agreement_ad_abuse_type <- tibble(
  CaseID          = c("1001", "1002", "1003", "1004", "1005", "1006"), 
  physical        = rep("Agree", 6), 
  sexual          = rep("Agree", 6), 
  emotional       = c("Agree", "Disagree", "Agree", "Agree", "Agree", "Agree"),
  neglect         = rep("Agree", 6), 
  abandonment     = c("Agree", "Agree", NA, "Agree", "Agree", "Agree"), 
  financial       = rep("Agree", 6), 
  selfneglect     = rep("Agree", 6), 
  TotalAgreement  = c(TRUE, FALSE, FALSE, TRUE, TRUE, TRUE), 
  AnyDisagreement = c(FALSE, TRUE, TRUE, FALSE, FALSE, FALSE)
)
```

Make table

```{r}
summary_agreement_ad_abuse_type_ft <- flextable(
  # Remove unneeded columns
  summary_agreement_ad_abuse_type %>% 
    select(
      CaseID, physical, sexual, emotional, neglect, 
      abandonment, financial, selfneglect
    )
  ) %>% 
  # Column width: Trial and error
  # Make a table and play with properties
  width(
    j = c(1:8), 
    width = c(0.98, 0.66, 0.56, 0.78, 0.71, 1.01, 0.71, 0.90)
  ) %>% 
  # Improve readability of column headers
  set_header_labels(CaseID = "Case Number", selfneglect = "Self Neglect") %>% 
  # Add title to top of table
  # Add a blank title line to top of table
  add_header_lines("") %>% 
  # Use compose to bold "Table #."
  compose(
    i = 1, part = "header",
    value = as_paragraph(
      as_chunk("Table 2. ", props = fp_text(bold = TRUE)),
      "Presence/absence of unanimous agreement for each abuse type by case number."
    ),
  ) %>%
  # Change font to times new roman
  font(fontname = "Times New Roman", part = "all") %>% 
  # Change background color of first column
  bg(j = 1, bg = "#E5E8E8", part = "body") %>% 
  # Center column headings
  align(i = 2, align = "center", part = "header") %>% 
  # Center body text
  align(align = "center", part = "body") %>% 
  # Conditionally format disagree to red
  color(i = ~ physical == "Disagree", j = c("CaseID", "physical"), color = "red") %>%
  color(i = ~ sexual == "Disagree", j = c("CaseID", "sexual"), color = "red") %>%
  color(i = ~ emotional == "Disagree", j = c("CaseID", "emotional"), color = "red") %>%
  color(i = ~ neglect == "Disagree", j = c("CaseID", "neglect"), color = "red") %>%
  color(i = ~ abandonment == "Disagree", j = c("CaseID", "abandonment"), color = "red") %>%
  color(i = ~ financial == "Disagree", j = c("CaseID", "financial"), color = "red") %>%
  color(i = ~ selfneglect == "Disagree", j = c("CaseID", "selfneglect"), color = "red")

# For checking
summary_agreement_ad_abuse_type_ft
```

<!--
Clean up and add session info
-->

```{r echo=FALSE}
rm(list = ls())
```


