# Appendix A: Terms and Definitions

<!--
Hidden Notes
-->


Need to sort and define:

Terms to add to R Notes:
* Probability. Look up
* Risk. 57, but look up.
* Incidence time. 58
* Censoring. 59
* Person time at risk. 60
* Incidence rate. 61
* Inverse time. I think it just means per year, per month, per week, etc. 61



```{r random-pic-from-url, echo=FALSE}
knitr::include_graphics("https://picsum.photos/200/300")
```

## ⭐️Overview

There are so many terms in epidemiology that are used inconsistently. This note is supposed to help keep track of them all.

## 🌎Useful websites

* https://stackoverflow.com/

## 📖Terms

### General Terms

<p class="note"> 🗒**Side Note:** At some point, you may want to improve the organization of this note and/or add some graphics. </p>

*Cohort*. A group of persons for whom membership is defined in a permanent fashion, or a population in which membership is determined by satisfying a set of defining events and so becomes permanent. An example of a cohort would be the members of the graduating class of a school in a given year... With this definition, the members of any cohort constitute a closed population along the time axis in which the chronologically last defining event (e.g., birth with Down syndrome or study recruitment) is taken as the index time. @Lash2021-rz

*Closed population*. A closed population adds no new members over time and loses members only to death. @Lash2021-rz

*Open population*. Open populations may gain members over time, through immigration or birth, or lose members over time through emigration or death. @Lash2021-rz Think of a population defined by a geographic area or political boundary.

### General statistics terms

*Odds*. 

*Proportion*.

*Percentage*.

*Ratio*.

### Measures of occurence

*Incident*. A single occurrence of an outcome of interest. We might also talk about the number of incident cases. 

*Incidence*. 

*Incidence proportion*. 

*Prevalence*. 

<!--
Clean up and add session info
-->

```{r echo=FALSE}
rm(list = ls())
```

```{r echo=FALSE}
sessionInfo()
```
