# Epidemiology Analysis Terms

<!--
Hidden comments placeholder
-->

```{r random-pic-from-url, echo=FALSE}
knitr::include_graphics("https://picsum.photos/200/300")
```

## ⭐️Overview

There are so many terms in epidemiology that are used inconsistently. This note is supposed to help keep track of them all.

## 🌎Useful websites

* https://stackoverflow.com/

## 📖Terms

### General Terms

<p class="note"> 🗒**Side Note:** At some point, you may want to improve the organization of this note and/or add some graphics. </p>

*Cohort*. A group of persons for whom membership is defined in a permanent fashion, or a population in which membership is determined by satisfying a set of defining events and so becomes permanent. An example of a cohort would be the members of the graduating class of a school in a given year... With this definition, the members of any cohort constitute a closed population along the time axis in which the chronologically last defining event (e.g., birth with Down syndrome or study recruitment) is taken as the index time. @Lash2021-rz

*Closed population*. A closed population adds no new members over time and loses members only to death. @Lash2021-rz

*Open population*. Open populations may gain members over time, through immigration or birth, or lose members over time through emigration or death. @Lash2021-rz Think of a population defined by a geographic area or political boundary.

### Basic quantities

*Odds*. 

*Proportion*.

*Percentage*.

*Ratio*.

<!--
Clean up and add session info
-->

```{r echo=FALSE}
rm(list = ls())
```

```{r echo=FALSE}
sessionInfo()
```
