# Contingency Tables

<!--
Don't forget to add code for converting from a contingency table back into raw data. See the useful websites below. 
-->

```{r echo=FALSE}
knitr::include_graphics("https://picsum.photos/200/300")
```


## ⭐️Overview

In Epi III (and epidemiology in general) we use a lot of contingency tables - especially 2x2 contingency tables. In this note, we play around with several different ways of creating, and working with, contingency tables. This can include converting them to a data frame format for some analyses. We explore some of the pros and cons of each of each method.


## 🌎Useful websites

* [Boston University](https://sphweb.bumc.bu.edu/otlt/mph-modules/ep/ep713_association/ep713_association_print.html)

* [Tables in R – A quick practical overview](https://cran.r-project.org/web/packages/DescTools/vignettes/TablesInR.pdf). By Andri Signorell.

* [Stack Overflow: Repeat each row of data.frame the number of times specified in a column](https://stackoverflow.com/questions/2894775/repeat-each-row-of-data-frame-the-number-of-times-specified-in-a-column)

* [Cookbook for R](http://www.cookbook-r.com/Manipulating_data/Converting_between_data_frames_and_contingency_tables/)

* [Stak Overflow: Convert contingency table to long data.frame](https://stackoverflow.com/questions/48330888/r-convert-contingency-table-to-long-data-frame)


## 📦Load packages

```{r message=FALSE}
library(dplyr, warn.conflicts = FALSE)
library(freqtables)
```


## Terminology

In the text below, we use the following terminology to distinguish three different structural representations of our data:

* `Data frame of observations`: A data frame where each row represents one observation (typically an individual person). 

```{r}
# Basic example
df <- tibble(
  medication = c(rep("No", 10), rep("Yes", 10)),
  fall = c(rep("No", 8), rep("Yes", 2), rep("No", 6), rep("Yes", 4))
) %>% 
print()
```

* `Contingency table`: A grid-like table with the categories of one variable (typically the exposure of interest) making up the rows and the categories of a second variable (typically the outcome of interest) making up the columns of the table. The table contains the number of observations with the particular combination of row and column values that intersect at each cell.

```{r}
# Basic example
table(df)
```

* `Frequency table`: A data frame of counts (and optionally, other relevant statistics), where each row represents a particular combination of values from two or more categorical variables.

```{r}
# Basic example
df %>% 
  count(medication, fall)
```

As the examples above illustrate, it's pretty easy and straightforward to go from a data frame of observations to either a contingency table or a frequency table. What is currently less straightforward (or at least less well documented) is to:

* Manually create a contingency table.

* Convert a contingency table into a frequency table.

* Convert a contingency table into a data frame of observations.

* Convert a frequency table into a contingency table. 

* Convert a frequency table into a data frame of observations.

We go through each of these operations below.


## Scenario

This scenario is borrowed from the [Boston University](https://sphweb.bumc.bu.edu/otlt/mph-modules/ep/ep713_association/ep713_association_print.html) website. 

Data Summary

Consider the following example regarding the management of Hodgkin lymphoma, a cancer of the lymphatic system. Years ago when a patient was diagnosed with Hodgkin Disease, they would frequently undergo a surgical procedure called a "staging laparotomy." The purpose of the staging laparotomy was to determine the extent to which the cancer had spread, because this was important information for determining the patient's prognosis and optimizing treatment. At times, the surgeons performing this procedure would also remove the patient's appendix, not because it was inflamed; it was done "incidentally" in order to ensure that the patient never had to worry about getting appendicitis. However, performing an appendectomy requires transecting it, and this has the potential to contaminate the abdomen and the wound edges with bacteria normally contained inside the appendix. Some surgeons felt that doing this "incidental appendectomy" did the patient a favor by ensuring that they would never get appendicitis, but others felt that it meant unnecessarily increasing the patient's risk of getting a post-operative wound infection by spreading around the bacteria that was once inside the appendix. To address this, the surgeons at a large hospital performed a retrospective cohort study. They began by going through the hospital's medical records to identify all subjects who had had a "staging laparotomy performed for Hodgkin." They then reviewed the medical record and looked at the operative report to determine whether the patient had an incidental appendectomy or not. They then reviewed the progress notes, the laboratory reports, the nurses notes, and the discharge summary to determine whether the patient had developed a wound infection during the week after surgery. The investigators reviewed the records of 210 patients who had undergone the staging procedure and found that 131 had also had an incidental appendectomy, while the other 79 had not. The data from that study are summarized in the table below. The numbers in the second and third columns indicate the number of subjects who did or did not develop a post-operative wound infection among those who had the incidental appendectomy (in the "Yes" row) and those who did not have the incidental appendectomy (in the "No" row). For example, the upper left cell indicates that seven of the subjects who had an incidental appendectomy (the exposure of interest) subsequently developed a wound infection. The upper right cell indicates that the other 124 subjects who had an incidental appendectomy did NOT develop a wound infection.

|                              |                 |                    |       |
|------------------------------|-----------------|--------------------|-------|
| Had incidental appendectomy? | Wound infection | No wound infection | Total |
| Yes                          | 7               | 124                | 131   |
| No                           | 1               | 78                 | 79    |
| Total                        | 8               | 202                | 210   |


## Manually creating each data structure

In this section, we manually create the incidental appendectomy data in all three structural representations (i.e., `Data frame of observations`, `Contingency table`, `Frequency table`).

### Data frame of observations

First, we can manually create tibble with one row for each person represented in the data above. Ordinarily, this is how the data would come to us. Then, we can use various different techniques -- some of which are demonstrated below -- to summarize the data as a 2x2 contingency table. In this case, we are working backwards from the data summary to the raw data just to show one way that it can be done. This isn't necessarily a good way to do it, however. Later, we will demonstrate more efficient and less error-prone ways to create raw data from summary tables. 

```{r}
df <- tibble(
  appendectomy = factor(c(rep("Yes", 7), rep("Yes", 124), "No", rep("No", 78))),
  infection    = factor(c(rep("Yes", 7), rep("No", 124), "Yes", rep("No", 78)))
) %>% 
print()
```

### Contingency tables

#### Matrix object

Create a contingency table as a **matrix** object. 

```{r}
matrix_ct <- matrix(
  c(a = 7, b = 124, c = 1, d = 78),
  ncol = 2,
  byrow = TRUE
)

# Add names to make the matrix more readable
rownames(matrix_ct) <- c("Appendectomy", "No Appendectomy")
colnames(matrix_ct) <- c("Infection", "No Infection")

matrix_ct
```

Adding summary rows/columns to the matrix.

```{r}
matrix_ct_margins <- cbind(matrix_ct, rowsum = rowSums(matrix_ct))
matrix_ct_margins <- rbind(matrix_ct_margins, colsum = colSums(matrix_ct_margins))
matrix_ct_margins
```

#### Table object

The table function "uses the cross-classifying factors to build a contingency table of the counts at each combination of factor levels." (https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/table). This is really R's built-in method for working with contingency tables. However, I haven't found a way to create a table object from scratch (i.e., without making a matrix or data frame first). Because this section is about manually creating each data structure, we won't say anything more about tables here. Later, we will use the `table()` function to convert matrices and data frames to table objects. 

#### Base R data frame with rownames

A third option for manually creating a contingency table is to start with a base R data frame with row names. 

```{r}
df_ct <- data.frame(
  Infection = c(7, 1),
  `No Infection` = c(124, 78)
)

# Add row names
rownames(df_ct) <- c("Appendectomy", "No Appendectomy")

df_ct
```

Add margins:

```{r}
df_ct_margins <- df_ct
df_ct_margins <- cbind(df_ct_margins, rowsum = rowSums(df_ct_margins))
df_ct_margins <- rbind(df_ct_margins, colsum = colSums(df_ct_margins))
df_ct_margins
```

As far as I can tell, there doesn't seem to be an advantage to creating a df from scratch over creating a matrix from scratch. 

### Frequency table

Unlike the other objects, this object won't have the standard contingency table layout. 

```{r}
freq_tbl <- tribble(
  ~appendectomy, ~infection, ~count,
  "Yes", "Yes", 7,
  "Yes", "No",  124,
  "No",  "Yes", 1,
  "No",  "No",  78
)

# Add margins
freq_tbl %>% 
  group_by(appendectomy) %>% 
  mutate(appendectomy_totals = sum(count)) %>% 
  group_by(infection) %>% 
  mutate(infection_totals = sum(count)) %>% 
  ungroup() %>% 
  mutate(margin_total = sum(count))
```


## Table objects

As stated above, the table object is really R's built-in method for working with contingency tables. However, I haven't found a way to create a table object from scratch (i.e., without making a matrix or data frame first). In this section, we briefly demonstrate how to create a table object from a matrix and from a data frame.

### Table object from matrix

First, we will create a table from a matrix object. 

```{r}
table_from_matrix <- as.table(matrix_ct)
table_from_matrix
```

```{r}
# Add margins
table_from_matrix_margins <- addmargins(table_from_matrix)
table_from_matrix_margins
```

Potential advantages to converting a matrix to a table object:   

* Adding margins in one step.

* Later, it's easier to convert a table object to a data frame of observations than it is to convert a matrix object to a data frame of observations. 

### Table object from df

Next, we will create a table from a data frame object.

```{r}
table_from_df <- table(df)
table_from_df
```

```{r}
# Add margins
table_from_df_margins <- addmargins(table_from_df)
table_from_df_margins
```

Potential advantages to converting a data frame contingency table to a table object:   

* Adding margins in one step.

* Later, it's easier to convert a table object to a data frame of observations than it is to convert a data frame contingency table to a data frame of observations. 


## Convert a contingency table into a frequency table.

### Matrix object

When starting with a matrix contingency table, the easiest way to convert it to a frequency table is to first convert it to a [table object][Table-objects] using `as.table()`. Then, we pass that result to `as.data.frame()` to create a frequency table. This solution comes from [Stack Overflow](https://stackoverflow.com/questions/48330888/r-convert-contingency-table-to-long-data-frame).

```{r}
matrix_ct %>% 
  as.table() %>% 
  as.data.frame()
```

### Table object from df

When starting with a [table object made from a df][Table-object-from-df], we only need to pass it to the `as.data.frame()` function. Notice that the result is slightly different than above. 

```{r}
df %>% 
  table() %>% 
  as.data.frame()
```

### Data frame contingency table

When starting from a [base R data frame contingency table][base-r-data-frame-with-rownames], there are a couple of options. [These two come from Stack Overflow](https://stackoverflow.com/questions/48330888/r-convert-contingency-table-to-long-data-frame). 

The first method is a bit convoluted, but uses only base R functions. 

```{r}
df_ct %>% 
  as.matrix() %>% 
  as.table() %>% 
  as.data.frame()
```

The second method is a `Tidyverse` solution. 

```{r}
df_ct %>% 
  tibble::rownames_to_column() %>% 
  tidyr::pivot_longer(c(Infection, No.Infection))
```


# 🔴Here down needs refinement

Also, do we want to create a section of functions that can do each of these things in a single step?

## Convert a contingency table into a data frame of observations.

This is a really common conversion to want to make. We often come across data that is already in a 2x2 table and decide that we want to convert it into a data frame of observations to experiment with it or for some statistical procedures (e.g., regression).

All of the methods I've found so far require creating a frequency table as an intermediate step. 

### Matrix object

```{r}
# Convert from data frame of counts to data frame of cases.
# `countcol` is the name of the column containing the counts
countsToCases <- function(x, countcol = "Freq") {
    # Get the row indices to pull from x
    idx <- rep.int(seq_len(nrow(x)), x[[countcol]])

    # Drop count column
    x[[countcol]] <- NULL

    # Get the rows from x
    x[idx, ]
}
```

```{r}
matrix_ct %>% 
  as.table() %>% 
  as.data.frame() %>% 
  countsToCases() %>% 
  tibble()
```

Works, but I don't like it.

### Table object {#table-from-matrix-to-df-obs}

Same as above, there are just fewer steps because you don't have to convert to a table object first.

```{r}
table_from_matrix %>% 
  as.data.frame() %>% 
  countsToCases() %>% 
  tibble()
```

```{r}
table_from_df %>% 
  as.data.frame() %>% 
  countsToCases() %>% 
  tibble()
```

Again, these work, but I don't like them. I guess I could just drop the row numbers. 


## Convert a frequency table into a contingency table. 

http://www.cookbook-r.com/Manipulating_data/Converting_between_data_frames_and_contingency_tables/

```{r}
xtabs(count ~ appendectomy + infection, data = freq_tbl)
```

Again, this works, but I don't love it for some reason. One reason is that "No" is above "Yes".


## Convert a frequency table into a data frame of observations.

We already saw how to do this [above](#table-from-matrix-to-df-obs) as an intermediate stepe between a contingency table and a data frame of observations. 

```{r}
freq_tbl %>% 
   countsToCases("count")
```


## Other convertions 

I'm not sure where to put this stuff yet. 

Create a base R data frame by coercing the matrix object to a data frame:

```{r}
df_from_matrix_margins <- as.data.frame(matrix_ct_margins)
df_from_matrix_margins
```

Coercing the table object to a data frame gives a different result:

```{r}
as.data.frame(table_from_df_margins)
```

At this point, I think it still makes sense to start with a matrix. You can easily convert a matrix to a data frame later if you need to. 


## Bottom line

When we open a open a book or journal article and see a 2x2 table, or when we just decide to create a 2x2 table from scratch to do some experimenting, it looks like it's best to start with a matrix and convert it to a table object. 

```{r}
matrix_ct <- matrix(
  c(a = 7, b = 124, c = 1, d = 78),
  ncol = 2,
  byrow = TRUE
)

# Add names to make the matrix more readable
rownames(matrix_ct) <- c("Appendectomy", "No Appendectomy")
colnames(matrix_ct) <- c("Infection", "No Infection")

matrix_ct
```

```{r}
table_from_matrix <- matrix_ct %>% 
  as.table() %>% 
  print()
```

Optionally add margins

```{r}
table_from_matrix %>% 
  addmargins()
```

We can condense all of this down into a function that makes it easy to do in one step.

[Insert function code here]

To then convert that contingency table to a data frame of observations (another common scenario), we use...

```{r}
table_from_matrix %>% 
  as.data.frame() %>% 
  countsToCases() %>% 
  tibble()
```

Again, we can condense all of this down into a function that makes it easy to do in one step.

[Insert function code here]


I still want to rewrite the contsToCases code for some reason. 

It comes from here: http://www.cookbook-r.com/Manipulating_data/Converting_between_data_frames_and_contingency_tables/

Check: https://stackoverflow.com/questions/2894775/repeat-each-row-of-data-frame-the-number-of-times-specified-in-a-column

And: https://stackoverflow.com/questions/48330888/r-convert-contingency-table-to-long-data-frame

For alternatives.


<!--
Clean up and add session info
-->

```{r echo=FALSE}
rm(list = ls())
```


